Concurrent Programming with CXL

### 2020-06-15 book/paper.tex
Latest

A good example of a resource with multiple units is the so-called \emph{bounded buffer}.
It is essentially
a queue implemented using a circular buffer of a certain length and two pointers:
one where items are inserted and one from where items are extracted.  If the
buffer is full, processes that want to add more items have to wait.  In the usual
formulation, processes that want to extract an item when the buffer is empty
also have to wait.
This problem is known as the ``Producer/Consumer Problem.''

@@

The code in Figure~\ref{fig:hw} is based on code developed by Herlihy and
Wing~\cite{HW}.  It uses two interlock instructions:
\begin{itemize}
\item \texttt{inc(p)}: atomically increments \texttt{\^{}p} and returns
the old value;
\item \texttt{swap(p)}: sets \texttt{\^{}p} to \texttt{()} and returns
the old value.
\end{itemize}

>>>>

A good example of a resource with multiple units is the
so-called \emph{bounded buffer} (aka \emph{ring buffer}).
It is essentially
a queue implemented using a circular buffer of a certain length and two pointers:
one where items are inserted and one from where items are extracted.  If the
buffer is full, processes that want to add more items have to wait.  In the usual
formulation, processes that want to extract an item when the buffer is empty
also have to wait.
This problem is known as the ``Producer/Consumer Problem.''

@@

The code in Figure~\ref{fig:hw} is based on code developed by Herlihy and
Wing~\cite{HW}.
The code is a ``proof of existence'' for non-blocking synchronization; it
is not necessarily practical.
There are two variables.  \texttt{items} is an unbounded array with each
entry initialized to \texttt{()}.  \texttt{back} is an index into the
array and points to the next slot where a new value is inserted.
The code uses two interlock instructions:
\begin{itemize}
\item \texttt{inc(p)}: atomically increments \texttt{\^{}p} and returns
the old value;
\item \texttt{swap(p)}: sets \texttt{\^{}p} to \texttt{()} and returns
the old value.
\end{itemize}

Method \texttt{produce(item)} uses \texttt{inc \&(back)} to allocate
the next available slot in the \texttt{items} array.
It stores the item as a singleton tuple.
Method \texttt{consume()} repeatedly scans the array, up to the
\texttt{back} index, trying to find an item to return.
To check an entry, it uses \texttt{swap()}
to atomically remove an item from a slot if there is one.
This way, if two or more processes attempt to extract an item from
the same slot, at most one will succeed.

There is no critical section.  If one process is executing instructions
very slowly, this does not negatively impact the other processes, as it
would with solutions based on critical sections.
On the contrary, it helps them because it creates less contention.

The solution is not practical because the \texttt{items} array must be
unbounded if an unbounded number of items may be produced.  Each slot
in the array is only used once, which is inefficient.  Also, the \texttt{inc}
and \texttt{swap} interlock instructions are not universally available
on existing hardware.
However, in the literature there are many examples of practical
non-blocking (aka \emph{wait-free}) synchronization algorithms~\cite{}.
